---
title: "Week 12 IP"
author: "Njoki Mbugua"
date: "7/1/2021"
output:
  html_document: default
  pdf_document: default
---

```{r}
library(knitr)
```

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Defining the Question
### a) Specifying the Question
The main objective of this project is to determine which individuals are most likely to click on ads in a blog post.

### b) Defining the Metric for Success
Exhaustively perform Exploratory Data Analysis (Univariate and bivariate analysis) to determine that enabling factors of individuals clicking on ads.

### c) Understanding the context
Working as a consultant Data science consultant to help the client identify which individuals are most likely to click on her ads. The entrepreneur has created an online cryptography course and would want to advertise it on her blog. She currently targets audiences originating from various countries. In the past, she ran ads to advertise a related course on the same blog and collected data in the process.

### d) Recording the experimental design
Importing and reading the data
Data Cleaning
Exploratory Data Analysis(Univariate and bivariate)
Conclusions and recommendations

### e) Data Relevance
The data columns in the data set include:Daily Time Spent on Site,Age,Area Income,Daily Internet Usage,Ad Topic Line,City,Male,Country,Time stamp and Clicked on Ad(http://bit.ly/IPAdvertisingData).

# Reading the data
```{r}

data <- read.csv(file.choose(),na.strings = "")

# printing the top six rows of the dataset
head(data)

```

```{r}
# printing the bottom rows of the dataset

tail(data, n=10)

```

```{r}
# checking the number of rows in the dataset
nrow(data)

#We have 1000 rows.
```

```{r}
# checking the number of columns
ncol(data)

#The dataset has 10 columns
```

```{r}
# checking structure of the dataset
str(data)
```

```{r}
# getting summary on the columns
summary(data)

```
# Data Cleaning
## Checking for missing values
```{r}
sum(is.na(data))

```
There are no missing values in the data set.

## Checking for duplicates
```{r}

# duplicated(data)
sum(duplicated(data))

# there are no duplicates in the dataset

```
There are no duplicates in the data set.
```{r}
install.packages("tidyverse", repos = 'http://cran.us.r-project.org')

library(tidyverse)
```

```{r}
library(dplyr)
```

```{r}
# Selecting numeric column
df <- select_if(data, is.numeric)             
df
```
## Checking for outliers

```{r}
# Using the boxplot to detect outliers in the numeric columns
boxplot(df)

```
Outliers were detected in the income column.


Listing the outliers in the income column
```{r}
# listing outliers
boxplot.stats(df$Area.Income)$out
```
The outliers were not removed since income levels vary extremely meaning that they are valid and important for analysis.

# Univariate analysis
## Measures of central tendency
```{R}
# Time spent on site
Daily.Time.Spent.on.Site.mean <- mean(data$Daily.Time.Spent.on.Site)
Daily.Time.Spent.on.Site.mean

Daily.Time.Spent.on.Site.median <-median(data$Daily.Time.Spent.on.Site)
Daily.Time.Spent.on.Site.median
Daily.Time.Spent.on.Site.sd <- sd(data$Daily.Time.Spent.on.Site)
Daily.Time.Spent.on.Site.sd
```
The mean daily time spent on site is 65.002
```{R}
# Income mean
##Area.Income.mean
```
The mean income is 55000.
```{R}
# Average age
##Age.mean
```
The average age is 36.
```{R}
# getting summary statistics for numeric variables
summary(df)
```
The mean age of people visiting the site is 36, max age is 61 and min age is 19.

The maximum area income is 79485 and min area income  is 13996 with mean area income being 55000. 

The minimum amount of time spent on the blog is 32.60 and maximum is 91.43 with a mean of 65.


The maximum daily internet usage on the website is 270 with a mean of 180 and a minimum of 104.

```{r}
install.packages("moments", repos = 'http://cran.us.r-project.org')
```

```{r}
library(moments)
```

```{R}
install.packages("ggplot2", repos = 'http://cran.us.r-project.org')

```

```{r}
library(ggplot2)
```

## Univariate graphical EDA
```{R}
#help("ggplot2")
# Histogram of age.
qplot(data=df, x=df$Age, colour=I("blue"))

```
Age is not normally distributed.
Most of the respondents are between the ages of 25 and 50.

```{R}
#library(ggplot2)
ggplot(data=df, aes(x = df$Age)) + 
    geom_bar(aes(fill=..count..), alpha=0.5, size=0.5,colour="blue",binwidth=60*5)

```


```{R}
# Histogram of Area income
qplot(data=df, x=df$Area.Income, colour=I("blue"))

```
Most People were earning between 50,000 and 70,000.


```{R}
# Histogram of Daily Internet Usage
qplot(data=df, x=df$Daily.Internet.Usage, colour=I("red"))


```
There is no particular pattern for the daily internet usage

```{R}
qplot(data=df, x=df$Daily.Time.Spent.on.Site, colour=I("red"))


```
Majority of the people spent about 60-80 minutes on the blog site.


# Bivariate Analysis
```{r}
# bargraph showing clicking of ads by gender
ggplot(data = df, aes(x = df$Male)) +  geom_bar(fill = "#0073C2FF", color = 'cyan')
```
Females are more likely to click on ads than males.

```{r}
# barplot showing count of how people clicked on ads

ggplot(data = df, aes(x = df$Clicked.on.Ad)) +  geom_bar(fill = "#0073C2FF", color = 'cyan')
  

```
There was an equal number of individuals who clicked on ad and those who did not.

```{R}
# relationship between income and clicking on ad

#qplot(data=data, x=data$Age,y=data$Area.Income, colour=data$Clicked.on.Ad, #size=I(3),shape=I(16),alpha=0.6, main = "Income vs clicking on ad")

ggplot(data = data) + 
	geom_point(aes(x = data$Age, y = data$Area.Income, color = data$Clicked.on.Ad), size = 3, alpha = 0.6)


```

People earning around 75k and below are more likely to click on ads. 

```{R}
# Scatterplot showing relationship between age, time spent on site and clicking on ads

ggplot(data = data) +
  geom_point(aes(x = data$Age, y = data$Daily.Time.Spent.on.Site, color = data$Clicked.on.Ad), size = 3, alpha = 0.6)
```
Most people that log in to the site are between the ages of 30-50.
The time spent by the respondents on site does not necessarily translate to the respondents clicking on ads.
Most people that click on adds are between the ages of 40-50.
..............................

```{R}
# Scatterplot showing relationship between internet usage, time spent on site and clicking on ads

ggplot(data = data) +
  geom_point(aes(x = data$Daily.Internet.Usage, y = data$Daily.Time.Spent.on.Site, color = data$Clicked.on.Ad), size = 3, alpha = 0.6)

ggplot(data, aes(x = `Daily.Internet.Usage`, y = `Daily.Time.Spent.on.Site`)) +
  geom_point() 
```
Those with less daily internet usage tend to click more on ads
Spending much time on the site does not equal to clicking on the ad.


```{R}
install.packages("corrplot",repos = 'http://cran.us.r-project.org')

```

```{R}
library(corrplot)
```


```{R}
# Creating a correlation plot for the numeric variables

corrplot(cor(df), method="color",addCoef.col = "black", 
         tl.col="black", tl.srt=45)

corrplot(cor(df), method = "circle") 
```
There is a strong negative correlation between clicking on ad and daily time spent on site.

A strong negative correlation is also observed between clicking on ad and daily internet usage.


# Conclusions and recommendations
1. Females are more likely to click on ads than males
2. Middle income earners are more likely to click on ads as opposed to other income groups
3. More internet usage and time spent on the site doesn't not translate to individuals clicking on ads


# Modeling
Since it's a binary classification problem, we use logistic regression model as the baseline model.
```{r}
head(df)


```
```{r}
str(df)
```

```{r}

```

```{r}

```





























